Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-11 shanghai.tex
Finished proof of Lemma 10.1, removed some references to line bundles

The functor $\CS(G) \to \CS(G(\bFq))$ defined by $(\gcs{L}, \mu, \phi) \mapsto (\gcs{L}_x = x^*\gcs{L}, \mu_{x,y} = (x \times y)^* \mu, x^*\phi)$ is easily seen to be an equivalence.  We will use this equivalence implicitly in proving the following proposition.

\begin{proposition}\label{prop:etale}
  If $G$ is a commutative \'etale group scheme over $\Fq$ and
  $G(\bFq)$ is finitely generated then there is a canonical
  isomorphism
  \[
  \CSiso{G} \longrightarrow \Hh^1(\Weil{\Fq}, G(\bFq)^*),
  \qquad \text{given by}\qquad
  [\cs{L}] \mapsto [\tau_\cs{L}],
  \]
  where $\tau_\cs{L}: \Weil{\Fq}\to \Hom_\text{grp}(G(\bFq),\EEx)$ is a cocycle such that
  \[
  \forall g\in G(\Fq), \qquad \tau_\cs{L}(\Frob{\Fq})(g) = \trFrob{\cs{L}}(g).
  \]
\end{proposition}

\begin{proof}
  To produce the isomorphism between $\CSiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ we must pick a global section of the line
  bundle $(\gcs{L},\mu)$, for each character sheaf $\cs{L}$ on $G$; 
  such a section exists by Lemma~\ref{lemma:section}.
  We will show that the group isomorphism does not depend on these
  choices.
%
  We use the section $s_\cs{L}$ to define a cocycle $z_\cs{L} \in Z^1(\Weil{\Fq},G(\bFq)^*)$
  as follows. Recall that $\cs{L}$ also determines a family of linear transformations
  $\phi_x : \gcs{L}_{\Frob{\Fq}(x)} \to \gcs{L}_x$ of one-dimensional $\EE$-vector
  spaces, so $\phi_x(s(\Frob{\Fq}(x)))$ is a scalar multiply of $s(x)$, for
  each $x\in G(\bFq)$; let $\tau_\cs{L}(\Frob{\Fq})(x)$ be that scalar, so
  \begin{equation}\label{t}
    \forall x\in G(\bFq),\qquad \phi_{x}(s(\Frob{\Fq}(x)) = \tau_\cs{L}(\Frob{\Fq})(x) s(x).
  \end{equation}
  Conditions \ref{cs.0} and \ref{cs.2}, together with \eqref{section}, guarantee that
  the scalar $z_\cs{L}(\Frob{k})(x)$ is non-zero, while condition~\ref{cs.1} forces
  $\tau_\cs{L}(\Frob{\Fq})(x+y) = \tau_\cs{L}(\Frob{\Fq})(x) \ \tau_\cs{L}(\Frob{\Fq})(y)$,
  for all $x,y \in G(\bFq)$.
  % \Clifton{The alternative to this approach is to change the
  %   definition of character sheaves to incorporate the notion of
  %   Galois sheaves. That might not be so bad, but would add another
  %   layer of complication, which we wouldn't use profitably, expect
  %   here.}

@@

  Although we used a global section $s_\cs{L}$ in the definition of
  $\tau_\cs{L}$, the class of $\tau_\cs{L}$ does not depend on this
  choice. To see why, let $s_1$ and $s_2$ be two global sections of
  the line bundle $(\gcs{L},\mu)$, in the sense above and let $\tau_1$
  and $\tau_2$ be the cocycles defined by $s_1$ and $s_2$, respectively,
  as above. Since, for each $x\in G(\bFq)$, $\gcs{L}_x$ is a
  one-dimensional $\EE$-vector space, there is a unique scalar
  $a(x) \in \EE$ such that $s_2(x) = a(x) s_1(x)$; moreover, $a(x)$ is
  non-zero, since $s_2(x)$ is non-zero. Then \eqref{section} and
  condition~\ref{cs.1} together imply $a(x+y) = a(x)a(y)$, so
  $a\in G(\bFq)^*$. Now consider the coboundary $\tau_0 : \Weil{\Fq} \to G(\bFq)^*$
  defined by $\tau_0(\Frob{\Fq}) = \,^\Frob{\Fq} a \cdot a^{-1}$. Then $\tau_2 = \tau_0 \tau_1$,
  so $\tau_2$ and $\tau_1$ are cohomologous. In this way we see that
  \begin{equation}\label{cohomologous}
    \obj \CS(G(\bFq)) \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*),
    \qquad \text{defined by} \qquad \cs{L} \mapsto [\tau_\cs{L}],
  \end{equation}
  is independent of the choice of global sections made above.

@@

\begin{lemma}\label{lemma:section}
 If $\cs{L}$ is a character sheaf then there is a non-trivial section of $\gcs{L}$ 
 compatible with $\mu$ in the sense that 
 \begin{equation}\label{section}
  \forall x,y \in G(\bFq), \qquad \mu_{x,y}(s(x+y)) = s(x)\otimes s(y).
 \end{equation}
\end{lemma}

\begin{proof}
First, however, we show that such global sections exist.  Since
  $G(\bFq)$ is finitely-generated abelian, it admits a decomposition
  $G(\bFq) \iso G(\bFq)_\text{free} \oplus G(\bFq)_\text{tors}$ such
  that $G(\bFq)_\text{tors}$ is finite.  Let $x_1, \ldots , x_r$ be a
  basis for the free part $G(\bFq)_\text{free}$ of $G(\bFq)$ and let
  $y_1, \ldots , y_t$ be a generators for the torsion part of
  $G(\bFq)$ corresponding to the decomposition
  $G(\bFq)_\text{tors} \iso \oplus_{j=1}^t \ZZ/m_j\ZZ$, with
  $m_j \vert m_{j+1}$. For each such $x_i$ pick $s(x_i)\in \gcs{L}_{x_i}$;
  for each such $y_j$ pick $s(y_j) \in \gcs{L}_{y_j}$; also set
  $s({\bar e}) \ceq 0 \in \gcs{L}_{\bar e}$. ....
  \todo{under construction}
  In this way we define a section
  $s \ceq s_\cs{L} : G(\bFq) \to \coprod_{z\in G(\bFq)} \gcs{L}_z$ of
  projection $\coprod_{z\in G(\bFq)} \gcs{L}_z \to G(\bFq)$ such that
  $\mu_{x,y}(s(x+y)) = s(x)\otimes s(y)$ for all $x,y \in G(\bFq)$.
  Under the equivalence at the beginning of this section, $s = s_\cs{L}$
  corresponds to a global section of the line bundle $(\gcs{L},\mu)$.

\Clifton{I'm going to remove all mention of line bundles from this paper.}
\end{proof}

>>>>

The functor $\CS(G) \to \CS(G(\bFq))$ defined by
$(\gcs{L}, \mu, \phi) \mapsto (\gcs{L}_x = x^*\gcs{L}, \mu_{x,y} = (x \times y)^* \mu, x^*\phi)$
is easily seen to be an equivalence.  We may now define a \cdef{global section}
of a character sheaf $\cs{L}$ on an \'etale group scheme $G$ as a choice $s(x) \in \gcs{L}_x$ for every
$x \in G(\bFq)$, satisfying
 \begin{equation}\label{section}
  \forall x,y \in G(\bFq), \qquad \mu_{x,y}(s(x+y)) = s(x)\otimes s(y).
 \end{equation}

\begin{lemma}\label{lemma:section}
 If $\cs{L}$ is a character sheaf on a commutative, \'etale group scheme $G$
 with $G(\bFq)$ finitely generated then nontrivial global sections exist.
\end{lemma}

\begin{proof}
  Since $G(\bFq)$ is a finitely-generated abelian group, it admits a decomposition
  \[
  G(\bFq) \iso \ZZ^r \oplus \bigoplus_{j=1}^t \ZZ / m_j \ZZ;
  \]
  let $x_1, \ldots, x_r, y_1, \ldots, y_t$ be a corresponding set of generators of $G(\bFq)$.
  We will define a section $s$ by choosing values on these generators and extending to all
  of $G(\bFq)$ through \eqref{section}.

  There is a unique isomorphism $\gcs{L}_0 \cong \EE$ under which $\mu_{0,0}$ corresponds
  to the multiplication map $\EE \otimes \EE \to \EE$.
  The condition $\mu_{0,0}(s(0)) = s(0) \otimes s(0)$ together with the non-triviality of $s$
  then forces $s(0) = 1$.

  The $\mu_{x,y}$ yield an isomorphism $\alpha_j : \gcs{L}_{y_j}^{\otimes m_j} \to \gcs{L}_0$.
  An initial nonzero choice of $s(y_j)$ need only be scaled by $\alpha_j(s(y_j))^{-1/m_j}$ in order to map to
  $1$ under this isomorphism.  Since compatibility under the $\alpha_j$ are the only constraints on the
  $s(x_j)$ and $s(y_j)$, global sections exist.
\end{proof}

We will use this equivalence implicitly in proving the following proposition.

\begin{proposition}\label{prop:etale}
  If $G$ is a commutative \'etale group scheme over $\Fq$ and
  $G(\bFq)$ is finitely generated then there is a canonical
  isomorphism
  \[
  \CSiso{G} \longrightarrow \Hh^1(\Weil{\Fq}, G(\bFq)^*),
  \qquad \text{given by}\qquad
  [\cs{L}] \mapsto [\tau_\cs{L}],
  \]
  where $\tau_\cs{L}: \Weil{\Fq}\to G(\bFq)^*$ is a cocycle such that
  \[
  \forall g\in G(\Fq), \qquad \tau_\cs{L}(\Frob{\Fq})(g) = \trFrob{\cs{L}}(g).
  \]
\end{proposition}

\begin{proof}
  To produce the isomorphism between $\CSiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ we must pick a global section of $\cs{L}$, for each character sheaf $\cs{L}$ on $G$; 
  such a section exists by Lemma~\ref{lemma:section}.
  We will show that the group isomorphism does not depend on these
  choices.
%
  We use the section $s_\cs{L}$ to define a cocycle $z_\cs{L} \in Z^1(\Weil{\Fq},G(\bFq)^*)$
  as follows. Recall that $\cs{L}$ also determines a family of linear transformations
  $\phi_x : \gcs{L}_{\Frob{\Fq}(x)} \to \gcs{L}_x$ of one-dimensional $\EE$-vector
  spaces, so $\phi_x(s(\Frob{\Fq}(x)))$ is a scalar multiply of $s(x)$, for
  each $x\in G(\bFq)$; let $\tau_\cs{L}(\Frob{\Fq})(x)$ be that scalar, so
  \begin{equation}\label{t}
    \forall x\in G(\bFq),\qquad \phi_{x}(s(\Frob{\Fq}(x)) = \tau_\cs{L}(\Frob{\Fq})(x) s(x).
  \end{equation}
  Conditions \ref{cs.0} and \ref{cs.2}, together with \eqref{section}, guarantee that
  the scalar $z_\cs{L}(\Frob{k})(x)$ is non-zero, while condition~\ref{cs.1} forces
  $\tau_\cs{L}(\Frob{\Fq})(x+y) = \tau_\cs{L}(\Frob{\Fq})(x) \ \tau_\cs{L}(\Frob{\Fq})(y)$,
  for all $x,y \in G(\bFq)$.
  % \Clifton{The alternative to this approach is to change the
  %   definition of character sheaves to incorporate the notion of
  %   Galois sheaves. That might not be so bad, but would add another
  %   layer of complication, which we wouldn't use profitably, expect
  %   here.}

@@

  Although we used a global section $s_\cs{L}$ in the definition of
  $\tau_\cs{L}$, the class of $\tau_\cs{L}$ does not depend on this
  choice. To see why, let $s_1$ and $s_2$ be two global sections of
 $\cs{L}$ and let $\tau_1$
  and $\tau_2$ be the cocycles defined by $s_1$ and $s_2$, respectively,
  as above. Since, for each $x\in G(\bFq)$, $\gcs{L}_x$ is a
  one-dimensional $\EE$-vector space, there is a unique scalar
  $a(x) \in \EE$ such that $s_2(x) = a(x) s_1(x)$; moreover, $a(x)$ is
  non-zero, since $s_2(x)$ is non-zero. Then \eqref{section} and
  condition~\ref{cs.1} together imply $a(x+y) = a(x)a(y)$, so
  $a\in G(\bFq)^*$. Now consider the coboundary $\tau_0 : \Weil{\Fq} \to G(\bFq)^*$
  defined by $\tau_0(\Frob{\Fq}) = \,^\Frob{\Fq} a \cdot a^{-1}$. Then $\tau_2 = \tau_0 \tau_1$,
  so $\tau_2$ and $\tau_1$ are cohomologous. In this way we see that
  \begin{equation}\label{cohomologous}
    \obj \CS(G(\bFq)) \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*),
    \qquad \text{defined by} \qquad \cs{L} \mapsto [\tau_\cs{L}],
  \end{equation}
  is independent of the choice of global sections made above.

@@
